Analysis

### 2023-02-08 Analysis.rmd
update manuscript, add data revision plots, add screenshots of the app, add citations

```{r}
forecasts <- fread("data/forecast-data.csv")
scores <- fread("data/all-scores-crowd-forecasts.csv") |>
    mutate(type_and_scale = factor(type_and_scale, 
                                 levels = c("Cases - natural", "Deaths - natural", 
                                            "Cases - sqrt", "Deaths - sqrt", 
                                            "Cases - log", "Deaths - log")))
weekly_truth <- fread("data/weekly-truth.csv")
daily_truth <- fread("data/daily-truth.csv")
ensemble_inclusions <- fread("data/ensemble-inclusions.csv")

time_start <- as.Date("2021-05-24")
time_stop <- as.Date("2021-08-16")

@@

```{r}
weekly_truth |>
  ggplot(aes(x = target_end_date, y = true_value)) +
  annotate("rect", xmin = time_start, xmax = time_stop, ymin = 0, ymax = Inf,
           fill = "darkolivegreen3", alpha = .1) + 
  geom_line() + 
  geom_point(size = 0.8, aes(color = status)) + 
  geom_bar(data = daily_truth |>
             mutate(true_value = true_value * 7), 
           stat = "identity", 
           fill = "skyblue3", alpha = 0.4) + 
  facet_wrap(~ target_type, 
             nrow = 2, scale = "free_y") + 
  scale_y_continuous(labels = label_fn) +
  theme_scoringutils() + 
  scale_colour_manual(values = c("red", "black")) + 
  labs(y = "Observed value", x = "Date", color = "Data status")
```

@@

available <- function(forecasts, filter, name) {
  forecasts |>
    filter(model %in% filter) |>
    avail_forecasts(by = c("forecast_date", "target_type"), 
                    collapse = c("horizon", "quantile", "target_end_date")) |>
    mutate(type = name)
}

## Number of forecasts
avail <- 
  available(forecasts, filter = models_direct, name = "Direct") |>
  rbind(available(forecasts, filter = models_rt, name = "Rt"))

plot_num_forecasters <- avail |>
  ggplot(aes(x = forecast_date, y = `Number forecasts`, 
             fill = type)) +
  geom_bar(stat = "identity", color = "white") + 
  facet_wrap(~ target_type) + 
  labs(x = "Forecast date", fill = "Forecast method") + 
  theme_scoringutils()

ggsave("output/figures/num-forecasters-week.png", plot = plot_num_forecasters, 
       width = 7, height = 3)

## Number of Hub ensemble members
plot_ensemble_members <- ensemble_inclusions |>
  mutate(prediction = 1, true_value = 1) |>
  avail_forecasts(by = c("forecast_date", "target_type"), 
                    collapse = c("horizon", "quantile", "target_end_date")) |>
  ggplot(aes(x = forecast_date, y = `Number forecasts`)) +
  geom_bar(stat = "identity", fill = "grey60", color = "white") + 
  facet_wrap(~ target_type) + 
  labs(x = "Forecast date", fill = "Forecast method") + 
  theme_scoringutils()

ggsave("output/figures/num-ensemble-members.png", plot = plot_ensemble_members, 
       width = 7, height = 3)

@@

>>>>

```{r}
forecasts <- fread("data/forecast-data.csv")
scores <- fread("data/all-scores-crowd-forecasts.csv") |>
    mutate(type_and_scale = factor(type_and_scale, 
                                 levels = c("Cases - natural", "Deaths - natural", 
                                            "Cases - sqrt", "Deaths - sqrt", 
                                            "Cases - log", "Deaths - log")))
weekly_truth_revised <- fread("data/weekly-truth.csv")
daily_truth_revised <- fread("data/daily-truth.csv")
weekly_truth <- fread("data/weekly-truth-original.csv")
daily_truth <- fread("data/daily-truth-original.csv")

ensemble_inclusions <- fread("data/ensemble-inclusions.csv")
data_revisions <- fread("data/data-revisions-deaths.csv")

time_start <- as.Date("2021-05-24")
time_stop <- as.Date("2021-08-16") + 4 * 7 - 2

@@

```{r}
plot_data <- function(weekly_truth, daily_truth) {
  weekly_truth |>
  ggplot(aes(x = target_end_date, y = true_value)) +
  annotate("rect", xmin = time_start, xmax = time_stop, ymin = 0, ymax = Inf,
           fill = "darkolivegreen3", alpha = .1) + 
  geom_line() + 
  geom_point(size = 0.8, aes(color = status)) + 
  geom_bar(data = daily_truth |>
             mutate(true_value = true_value * 7), 
           stat = "identity", 
           fill = "skyblue3", alpha = 0.4) + 
  facet_wrap(~ target_type, 
             nrow = 2, scale = "free_y") + 
  scale_y_continuous(labels = label_fn) +
  theme_scoringutils() + 
  scale_colour_manual(values = c("red", "black")) + 
  labs(y = "Observed value", x = "Date", color = "Data status")
}

plot_data_revised <- plot_data(weekly_truth_revised, daily_truth_revised)
plot_data <- plot_data(weekly_truth, daily_truth)

ggsave("output/figures/plot-data-revised.png", plot = plot_data_revised, 
       width = 7, height = 5)

ggsave("output/figures/plot-data.png", plot = plot_data, 
       width = 7, height = 3.5)

Show data revisions

plot_weekly_revisions <- weekly_truth |>
  rbind(weekly_truth_revised) |>
  filter(target_type == "Deaths") |>
  select(target_end_date, true_value, target_type) |>
  group_by(target_end_date, target_type) |>
  mutate(diff = c(NA, diff(true_value))) |>
  filter(!is.na(diff)) |>
  ggplot(aes(x = target_end_date, y = diff)) + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80") + 
  geom_line() + 
  geom_point(size = 0.7) + 
  facet_wrap(~ target_type, ncol = 1, scale = "free_y") +
  labs(y = "Difference in data", x = "Date") + 
  coord_cartesian(ylim = c(-500, 500)) + 
  theme_scoringutils()

ggsave("output/figures/data-revisions.png", plot = plot_weekly_revisions, 
       width = 7, height = 2.5)

@@

```{r}
available <- function(forecasts, filter, name) {
  forecasts |>
    filter(model %in% filter) |>
    avail_forecasts(by = c("forecast_date", "target_type"), 
                    collapse = c("horizon", "quantile", "target_end_date")) |>
    mutate(type = name)
}

## Number of forecasts and ensemble members
avail <- 
  available(forecasts, filter = models_direct, name = "Direct") |>
  rbind(available(forecasts, filter = models_rt, name = "Rt")) |>
  mutate(group = "Human forecast")

avail_ensemble <- ensemble_inclusions |>
  mutate(prediction = 1, true_value = 1) |>
  avail_forecasts(by = c("forecast_date", "target_type"), 
                    collapse = c("horizon", "quantile", "target_end_date")) |>
  mutate(group = "Forecast Hub")

plot_number_forecasters <- avail |>
  pivot_wider(names_from = type, values_from = `Number forecasts`) |>
  mutate(Rt = Direct + Rt) |>
  pivot_longer(cols = c(Direct, Rt), names_to = "type",
               values_to = "Number forecasts") |>
  as.data.table() |>
  rbind(avail_ensemble, fill = TRUE) |>
  ggplot(aes(x = forecast_date, y = `Number forecasts`, 
             group = group,
             fill = group, 
             alpha = type)) +
  geom_bar(stat = "identity", 
           color = "white",
           linewidth = 0.2,
           width = 5,
           position = position_dodge(width = 5)) + 
  facet_wrap(~ target_type) + 
  scale_alpha_manual(values = c(1, 0.5), na.translate = FALSE) + 
  labs(x = "Forecast date", fill = "Forecast type", 
       alpha = "Forecast method") + 
  theme_scoringutils() +
  theme(legend.position = "bottom", 
        legend.box="vertical", legend.margin = margin())

ggsave("output/figures/num-forecasters.png", plot = plot_number_forecasters, 
       width = 7, height = 3.5)

@@

```{r}
forecasts$forecast_date |>
  unique()

forecasts$target_end_date |>
  unique()
```
